---
title: "South Africa COVID-19 Data Validation & Quality Checks-Post June 2022 2022 indicators"
author: "Vanessa Da Costa"
date: "last updated 8/19/2022"
output: html_document 
---

##Purpose:
This code was developed to check the quality of USAID South Africa's COVID-19 indicators
##Checks valid for July 2022 onwards & in this R script: 
Check 1, Check 2, Check 3, Check 4, Check 5, Check 6, Check 8, Check 9a, Check 10, Check 11, Check 12, Check 13

##Step 1: Installing Packages (ONE-TIME STEP)
Install these packages the first time you run R on your computer. You do not need to run these everytime you open R.

```{r}
install.packages("devtools")
install.packages("googledrive")
install.packages("gargle")
install.packages("googlesheets4")
devtools::install_github("USAID-OHA-SI/glamr", build_vignettes = TRUE)
vignette("credential-management", package = "glamr")
install.packages("validate")
install.packages("flextable")
install.packages("plyr")
install.packages("rmarkdown")
install.packages("stringr")
install.packages("dplyr")
install.packages("tidyverse")
install.packages("tidyr")
install.packages("readr")
install.packages("readxl")
install.packages("glamr")
install.packages("purrr")
install.packages("data.table")
install.packages("splitstackshape")
install.packages("gophr")
install.packages("keyring")
install.packages("fs")
install.packages("readxl")
install.packages("lubridate")
install.packages("glue")
install.packages("validate")
install.packages("plyr")
install.packages("here")

```

##Step 2: Running Libraries (EVERY TIME STEP)
Run these packages everytime you open R. The code is depedent on these pacakges.

```{r}
library(rmarkdown)
library(stringr)
library(dplyr)
library(tidyverse)
library(tidyr)
library(readr)
library(readxl)
library(glamr)
library(purrr)
library(data.table)
library(splitstackshape)
library(gophr)
library(keyring)
library(fs)
library(readxl)
library(lubridate)
library(glue)
library(validate)
library(plyr)
library(here)
library(googledrive)
library(gargle)
library(googlesheets4)




```

##Step 3: Authenticate Google Drive (ONE-TIME STEP)
```{r}
set_email("vdacosta@usaid.gov") #CHANGE TO YOUR EMAIL ADDRESS
load_secrets()
#after this step copy and paste into the R profile script and save
#now stored so don't have to authenticate
drive_auth()
gs4_auth()

```
##Step 4: Read in datasets using google drive (EVERY TIME STEP)
This section pulls from the live google sheets that partners update (files must be google sheets, cannot be XLSX file) 
```{r}
#Google Sheet (must be google sheet, cannot be XLSX file) 
#This section pulls from the live google sheets that partners update
ANOVA<-read_sheet(as_sheets_id('1d4O3gZwdD1qglk1FW442sYnoHQ9SBH6V60s09kEQrjs'), sheet = "Tracker v3", guess_max = 10000)

GUIDEHOUSE<-read_sheet(as_sheets_id('1As-AVLcqhduKL0jA-c92D0vlO0gh1wIuumfMzYwNzo0'), sheet = "Tracker v3", guess_max = 10000)

GETF_NEXTMILE<-read_sheet(as_sheets_id('1WHdEzsyelf20N60JoP2ascrKQynDvaOEnfV3SRVcdBA'), sheet = "Tracker v3", guess_max = 10000)

ADAPT_RTC <-read_sheet(as_sheets_id('1k2yDUZk49SRjTq3ceeXzogUw1TnMszjdXGtYJFwW060'), sheet = "RTC_v3", guess_max = 10000)
     
ADAPT_Aurum <-read_sheet(as_sheets_id('1k2yDUZk49SRjTq3ceeXzogUw1TnMszjdXGtYJFwW060'), sheet = "Aurum_v3", guess_max = 10000)

ADAPT_FPD <-read_sheet(as_sheets_id('1k2yDUZk49SRjTq3ceeXzogUw1TnMszjdXGtYJFwW060'), sheet="FPD_v3", guess_max = 10000)

ADAPT_Stellenbosch <-read_sheet(as_sheets_id('1k2yDUZk49SRjTq3ceeXzogUw1TnMszjdXGtYJFwW060'), sheet="Stellenbosch_v3", guess_max = 10000)

ADAPT_Genesis <-read_sheet(as_sheets_id('1k2yDUZk49SRjTq3ceeXzogUw1TnMszjdXGtYJFwW060'), sheet="Genesis_v3", guess_max = 10000)

ADAPT_INTRA <-read_sheet(as_sheets_id('1k2yDUZk49SRjTq3ceeXzogUw1TnMszjdXGtYJFwW060'), sheet="IntraHealth_v3", guess_max = 10000)

ADAPT_IHPS <-read_sheet(as_sheets_id('1k2yDUZk49SRjTq3ceeXzogUw1TnMszjdXGtYJFwW060'), sheet="IHPS_v3", guess_max = 10000)

ADAPT_Qode <-read_sheet(as_sheets_id('1k2yDUZk49SRjTq3ceeXzogUw1TnMszjdXGtYJFwW060'), sheet="Qode_v3", guess_max = 10000)

ADAPT_ReAction <-read_sheet(as_sheets_id('1k2yDUZk49SRjTq3ceeXzogUw1TnMszjdXGtYJFwW060'), sheet="ReAction!_v3", guess_max = 10000) 


ADAPT <-bind_rows(ADAPT_RTC,ADAPT_Aurum, ADAPT_FPD,ADAPT_Genesis,ADAPT_INTRA,
    ADAPT_IHPS,ADAPT_Qode,ADAPT_ReAction,ADAPT_Stellenbosch)


PARTNER_TOOLS <-bind_rows(ANOVA,GUIDEHOUSE, GETF_NEXTMILE, ADAPT) 



```

##Step 5: Restructure dataset so it's long and dates are in correct format
```{r}
df2<- PARTNER_TOOLS %>%  tidyr::gather(`Reporting Date`,`Value`, `31-Jul-22`:`31-Dec-23`) %>% 
  dplyr::mutate(`Reporting Date` =as.character(as.Date(`Reporting Date`, "%d-%b-%y"),"%d-%b-%y"))%>% 
  dplyr::mutate(`Reporting Date` = as.Date(`Reporting Date`,"%d-%b-%y")) %>% 
  #created unique ID so row with errors can be easily identified
  dplyr::mutate(uniqueid=paste(Partner, OrgUnit, `Dataelement Code`, `disagg category`, `Disaggregate`, `Reporting Date`, sep= "_")) %>% 
  dplyr::filter(`Reporting Date` == "2022-07-31" |`Reporting Date` == "2022-06-30") #update this to current and previous reporting month

```

##Step 6: Create data frames for each Level 1 Check and bind together
```{r}
#Dataset to check missing
check1<- df2 %>%
  dplyr::filter(`Reporting Date` == "2022-07-31")%>%  #UPDATE DATE
  dplyr:: mutate(check = ifelse(is.na(Value), "Value is NA", "NO ISSUE")) %>% 
  dplyr::filter(`check`== "Value is NA") %>% 
  tidyr::unnest(Value) %>% 
  dplyr::mutate(`Value`= as.numeric(`Value`)) %>%
  dplyr::mutate(`check_guidance`="Review the 'Value' column") %>% 
  dplyr::mutate(`check_num`="check1")

#Dataset to check 0
check2<- df2 %>% 
  dplyr::filter(`Reporting Date` == "2022-07-31")%>%  #UPDATE DATE
  dplyr::mutate(`Value`= as.character(`Value`)) %>%
  tidyr::unnest(Value) %>% 
  dplyr::mutate(`Value`= as.numeric(`Value`)) %>%
  dplyr::mutate(check = ifelse(`Value`==0, "VALUE IS 0", "NO ISSUE")) %>% 
  dplyr::filter(`check`== "VALUE IS 0") %>%
  dplyr::mutate(`Value`= as.numeric(`Value`)) %>%
  dplyr::mutate(`check_guidance`="Review the 'Value' column") %>% 
  dplyr::mutate(`check_num`="check2")

#Bind level 1 checks into 1 dataset
Level1 <-  dplyr::bind_rows(check1, check2)

```

##Step 7: Create data frames for each Level 2 Check
```{r}

#Dataset to check Unknown Sex == Male + Female
check3<- df2 %>%
  dplyr::filter(`Reporting Date` == "2022-07-31")%>%  #UPDATE DATE
  dplyr::filter(`Dataelement Code`== "CV.1.3-3" | `Dataelement Code`== "CV.1.4-6" | `Dataelement Code`== "CV.1.4-7" | `Dataelement Code`==  "CV.1.4-8" | `Dataelement Code`==  "CV.2.1-12"| `Dataelement Code`==  "CV.2.2-13"| `Dataelement Code`==  "CV.2.3-15"| `Dataelement Code`==  "CV.2.4-18"| `Dataelement Code`==  "CV.2.5-20"| `Dataelement Code`==  "CV.1.9-1" | `Dataelement Code`==  "CV.1.9-2"| `Dataelement Code`==  "CV.1.9-3")   %>%
  dplyr::filter (`disagg category`== "Sex") %>% 
  tidyr::unnest(Value) %>% 
  dplyr::mutate(`Value`= as.numeric(`Value`)) %>%
  dplyr::mutate(id3=paste(Partner, OrgUnit, `Dataelement Code`, `Reporting Date`, sep= "_"))%>%
  dplyr::select(-c(uniqueid)) %>% 
  dplyr::group_by(id3,`Disaggregate`) %>% 
  dplyr::mutate(`Value`= as.numeric(`Value`)) %>%
  tidyr::spread(key=Disaggregate, value=Value) %>% 
  dplyr::mutate(`Reporting Date`= as.character(`Reporting Date`)) %>% 
  dplyr::mutate(`Female`= as.numeric(`Female`)) %>% 
  dplyr::mutate(`Male`= as.numeric(`Male`)) %>%
  dplyr::mutate(`Unknown sex`= as.numeric(`Unknown sex`)) %>%
  dplyr::mutate_at( dplyr::vars("Female"), ~ tidyr::replace_na(.,0)) %>% 
  dplyr::mutate_at( dplyr::vars("Male"), ~   tidyr::replace_na(.,0)) %>% 
  dplyr::mutate_at( dplyr::vars("Unknown sex"), ~tidyr::replace_na(.,0)) %>% 
  dplyr::mutate(F_M_Total= Female + Male) %>% 
  dplyr::mutate(check = ifelse(`Unknown sex`==F_M_Total &`Unknown sex` >0 , "Unknown Sex == Male + Female", "NO ISSUE")) %>% 
 dplyr::filter(`check`== "Unknown Sex == Male + Female") %>% 
  dplyr::mutate(`check_guidance`="Review the 'Unknown Sex'& 'F_M_Total' columns") %>% 
  dplyr::mutate(`check_num`="check3") %>%
  mutate(`Reporting Date` =as.Date(`Reporting Date`, origin="1899-12-30")) %>%  
  dplyr::rename(uniqueid= id3)

#Dataset to check if Sex <> Vaccine brand
check4<- df2 %>% 
  dplyr::filter(`Reporting Date` == "2022-07-31")%>%  #UPDATE DATE
  dplyr::filter(`Dataelement Code`== "CV.1.4-6" | `Dataelement Code`== "CV.1.4-7" | `Dataelement Code`==  "CV.1.4-8" | `Dataelement Code`==  "CV.1.9-1" | `Dataelement Code`==  "CV.1.9-2"| `Dataelement Code`==  "CV.1.9-3")   %>% 
  tidyr::unnest(Value) %>% 
  dplyr::mutate(`Value`= as.numeric(`Value`)) %>%
  dplyr::mutate(id4=paste(Partner, OrgUnit, `Dataelement Code`, `Reporting Date`,`disagg category`, sep= "_")) %>% 
  dplyr::select(-c(Disaggregate,uniqueid)) %>% 
  dplyr::mutate(`Value`= as.numeric(`Value`)) %>%
  dplyr::group_by(id4) %>% 
  dplyr::mutate(`Value`= sum(`Value`)) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(id4_2=paste(Partner, OrgUnit, `Dataelement Code`, `Reporting Date`, sep= "_")) %>%
  dplyr::select(-c(id4)) %>% 
  dplyr::distinct() %>% 
  tidyr::spread(key=`disagg category`, value=Value) %>% 
  dplyr::mutate_at(vars("Vaccine brand"), ~replace_na(.,0)) %>% 
  dplyr::mutate_at(vars("Sex"), ~replace_na(.,0)) %>% 
  dplyr::mutate(`Vaccine brand`= as.numeric(`Vaccine brand`)) %>% 
  dplyr::mutate(`Sex`= as.numeric(`Sex`)) %>% 
  dplyr:: mutate(check = ifelse(`Sex`==`Vaccine brand`, "NO ISSUE", "Sex <> Vaccine brand")) %>%
  dplyr::filter(`check`== "Sex <> Vaccine brand") %>% #should the difference be +/- one because right now all errors are only a difference of 1
    dplyr::mutate(`check_guidance`="Review the 'Sex'& 'Vaccine brand' columns") %>% 
  dplyr::mutate(`check_num`="check4") %>% 
  dplyr::rename(uniqueid= id4_2)


#Dataset to check if same value reported for two or more of the disaggregate options
check5 <-df2 %>% 
    dplyr::filter(`Reporting Date` == "2022-07-31")%>%  #UPDATE DATE
   dplyr::filter(`Dataelement Code`== "CV.1.3-3" |`Dataelement Code`== "CV.1.4-5" | `Dataelement Code`== "CV.1.4-6" | `Dataelement Code`== "CV.1.4-7" | `Dataelement Code`==  "CV.1.4-8" | `Dataelement Code`==  "CV.2.1-11"|`Dataelement Code`==  "CV.2.1-12"| `Dataelement Code`==  "CV.2.2-13"| `Dataelement Code`==  "CV.2.3-15"| `Dataelement Code`==  "CV.2.4-18"| `Dataelement Code`==  "CV.2.5-20"|`Dataelement Code`==  "CV.2.6-22"| `Dataelement Code`==  "CV.1.9-1" | `Dataelement Code`==  "CV.1.9-2"| `Dataelement Code`==  "CV.1.9-3")    %>%  
   dplyr::mutate(id5=paste(Partner, OrgUnit, `Dataelement Code`, `disagg category`, `Reporting Date`, sep= "_")) %>% 
  dplyr::mutate(`Value`= as.character(`Value`)) %>%
  group_by(id5, Value) %>% 
  dplyr::mutate(
    count = row_number(), # Counts rows by group
    check = count > 1 & Value >"0"    # TRUE if there is more than one row per group & if value doesn't equal 0
  ) %>% 
    dplyr::mutate(`check`= as.character(`check`)) %>%
  mutate(check = recode(check, `TRUE` = 'Value reported in 2+ Disaggregate options', `FALSE` = 'NO ISSUE' )) %>% 
  group_by(id5) %>% 
  dplyr::filter(count== max(count)) %>% 
  distinct(id5,count, .keep_all = TRUE) %>%
  filter(`check`== "Value reported in 2+ Disaggregate options") %>% 
  dplyr::select(-c(uniqueid, count)) %>% 
  dplyr::mutate(`check_guidance`="Review the 'Value' & 'Disaggregate' column- check the other disaggregates in review for corresponding duplicate") %>% 
  dplyr::mutate(`check_num`="check5") %>% 
  dplyr::mutate(`Value`= as.character(`Value`)) %>%
  dplyr::filter(Value != "NULL") %>% 
  dplyr::filter(Value != "NA") %>% 
  dplyr::rename(uniqueid= id5)


#Dataset to check if same value reported for two or more districts or provinces
check6 <-df2 %>% 
    dplyr::filter(`Reporting Date` == "2022-07-31")%>%  #UPDATE DATE
    dplyr::filter(`Dataelement Code`== "CV.1.2-2" |`Dataelement Code`== "CV.1.3-3" |`Dataelement Code`== "CV.1.3-4" |`Dataelement Code`== "CV.1.4-5" | `Dataelement Code`== "CV.1.4-6" | `Dataelement Code`== "CV.1.4-7" | `Dataelement Code`==  "CV.1.4-8"| `Dataelement Code`==  "CV.2.2-13"| `Dataelement Code`==  "CV.2.3-15"|  `Dataelement Code`==  "CV.2.4-17"| `Dataelement Code`==  "CV.2.4-18"|  `Dataelement Code`==  "CV.2.5-19" |`Dataelement Code`==  "CV.2.5-20"|`Dataelement Code`==  "CV.2.5-21"| `Dataelement Code`==  "CV.1.9-1" | `Dataelement Code`==  "CV.1.9-2"| `Dataelement Code`==  "CV.1.9-3") %>%   
   dplyr::mutate(id6=paste(`Dataelement Code`, `disagg category`,`Disaggregate`, `Reporting Date`, sep= "_")) %>% 
  dplyr::group_by(id6, Value) %>% 
  dplyr::mutate(
    count = dplyr::row_number(), # Counts rows by group
    check   = count > 1 & Value >"0"   # TRUE if there is more than one row per group
  ) %>%
  dplyr::mutate(`check`= as.character(`check`)) %>%
  dplyr::mutate(check = dplyr::recode(check, `TRUE` = 'Value reported in 2+ District or Province options', `FALSE` = 'NO ISSUE' )) %>% 
  dplyr::group_by(id6) %>% 
  dplyr::filter(count== max(count)) %>% 
  dplyr::distinct(id6,count, .keep_all = TRUE) %>%
  dplyr::filter(`check`== "Value reported in 2+ District or Province options") %>% 
  dplyr::select(-c(uniqueid, count)) %>% 
    dplyr::mutate(`check_guidance`="Review the 'Value' & 'OrgUnit' column- check the other OrgUnit in review for corresponding duplicate") %>% 
  dplyr::mutate(`check_num`="check6") %>% 
  dplyr::mutate(`Value`= as.character(`Value`)) %>%
  dplyr::filter(Value != "NULL") %>% 
  dplyr::filter(Value != "NA") %>% 
  dplyr::rename(uniqueid= id6)


#Dataset to check if same value reported as previous month
check8 <-df2 %>% 
  dplyr::mutate(id8=paste(Partner, OrgUnit, `Dataelement Code`, `disagg category`, `Disaggregate`, sep= "_"))%>%
  dplyr::select(-c(uniqueid)) %>%
  dplyr::mutate(`Value`= as.character(`Value`)) %>%
  tidyr::unnest(Value) %>% 
  dplyr::mutate(`Value`= as.numeric(`Value`)) %>%
  dplyr::arrange(id8, `Reporting Date`) %>% 
  dplyr::group_by(id8) %>% 
  #create a column for previously reported value
  dplyr::mutate(value_prev_month = dplyr::lag(Value, n =1, default=NA)) %>% 
# filter where the current and previous values are the same
  dplyr::mutate(check = ifelse(value_prev_month==Value & Value >0, "Same Value Reported as Previous Month", "NO ISSUE")) %>% 
  dplyr::filter(`check`== "Same Value Reported as Previous Month") %>%
  dplyr::mutate(`check_guidance`="Review the 'Value' & 'value_prev_month' columns") %>% 
  dplyr::mutate(`check_num`="check8") %>% 
  dplyr::mutate(`Value`= as.character(`Value`)) %>%
  dplyr::mutate(`value_prev_month`= as.character(`value_prev_month`)) %>%
  dplyr::rename(uniqueid= id8)

  
#Dataset to check if value changed from previous month(numeric)
check9a <-df2 %>% 
    dplyr::filter(`Dataelement Code`== "CV.1.3-3" |`Dataelement Code`==  "CV.2.1-12" |`Dataelement Code`==  "CV.2.2-13"| `Dataelement Code`==  "CV.2.3-15"| `Dataelement Code`==  "CV.2.5-20") %>% 
  dplyr::mutate(id9a=paste(Partner, OrgUnit, `Dataelement Code`, `disagg category`, `Disaggregate`, sep= "_"))%>%
  dplyr::mutate(`Value`= as.character(`Value`)) %>%
  tidyr::unnest(Value) %>% 
  dplyr::mutate(`Value`= as.numeric(`Value`)) %>%
  dplyr::select(-c(uniqueid)) %>%
  dplyr::mutate(`Value`= as.numeric(`Value`)) %>%
  dplyr::arrange(id9a, `Reporting Date`) %>% 
  dplyr::group_by(id9a) %>% 
  #create a column for previously reported value
  dplyr::mutate(value_prev_month = dplyr::lag(Value, n =1, default=NA)) %>% 
# filter where the current and previous values change
  dplyr::mutate(check = ifelse(value_prev_month!= Value & Value >0, "Change from Previous Month", "NO ISSUE")) %>% 
  dplyr::filter(`check`== "Change from Previous Month") %>% 
  dplyr::mutate(`check_guidance`="Review the 'Value' & 'value_prev_month' columns") %>% 
  dplyr::mutate(`check_num`="check9a") %>% 
  dplyr::mutate(`Value`= as.character(`Value`)) %>%
  dplyr::mutate(`value_prev_month`= as.character(`value_prev_month`)) %>%
  dplyr::rename(uniqueid= id9a)


#Dataset to check if value changed from +/- 50% from previous reported value
check10 <-df2 %>% 
  dplyr::mutate(id10=paste(Partner, OrgUnit, `Dataelement Code`, `disagg category`, `Disaggregate`, sep= "_"))%>%
   dplyr::mutate(`Value`= as.character(`Value`)) %>%
  tidyr::unnest(Value) %>% 
  dplyr::mutate(`Value`= as.numeric(`Value`)) %>%
  dplyr::select(-c(uniqueid)) %>%
  dplyr::mutate(`Value`= as.numeric(`Value`)) %>%
  dplyr::arrange(id10, `Reporting Date`) %>% 
  dplyr::group_by(id10) %>% 
  #create a column for previously reported value
  dplyr::mutate(value_prev_month = dplyr::lag(Value, n =1, default=NA)) %>% 
  dplyr::mutate(Percent_change =(Value- value_prev_month)/value_prev_month) %>% 
# filter where the current and previous values change
  dplyr::mutate(check = ifelse(Percent_change>0.5  | Percent_change< -0.5, "+/- 50% Change from Previous Month", "NO ISSUE")) %>% 
  dplyr::filter(`check`== "+/- 50% Change from Previous Month") %>%
  dplyr::mutate(`check_guidance`="Review the'Value'm 'value_prev_month', & 'Percent_change' columns") %>% 
  dplyr::mutate(`check_num`="check10") %>% 
  dplyr::mutate(`Value`= as.character(`Value`)) %>%
  dplyr::mutate(`value_prev_month`= as.character(`value_prev_month`)) %>%
  dplyr::rename(uniqueid= id10) %>% 
  dplyr::mutate(`check`= as.character(`check`))


#Dataset to check if Sex <> Training
check11<- df2 %>% 
  dplyr::filter(`Reporting Date` == "2022-07-31")%>%  #UPDATE DATE
  dplyr::filter(`Dataelement Code`==  "CV.2.5-20") %>% 
  tidyr::unnest(Value) %>% 
  dplyr::mutate(`Value`= as.numeric(`Value`)) %>%
  dplyr::mutate(id11=paste(Partner, OrgUnit, `Dataelement Code`, `Reporting Date`,`disagg category`, sep= "_")) %>% 
  dplyr::select(-c(Disaggregate,uniqueid)) %>% 
  dplyr::mutate_at(vars("Value"), ~replace_na(.,0)) %>% 
  dplyr::mutate(`Value`= as.numeric(`Value`)) %>%
  dplyr::group_by(id11) %>% 
  dplyr::mutate(`Value`= sum(`Value`)) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(id11_2=paste(Partner, OrgUnit, `Dataelement Code`, `Reporting Date`, sep= "_")) %>%
  dplyr::select(-c(id11)) %>% 
  dplyr::distinct() %>% 
  tidyr::spread(key=`disagg category`, value=Value) %>% 
  dplyr::mutate_at(vars("Training"), ~replace_na(.,0)) %>% 
  dplyr::mutate_at(vars("Sex"), ~replace_na(.,0)) %>% 
  dplyr::mutate(`Training`= as.numeric(`Training`)) %>% 
  dplyr::mutate(`Sex`= as.numeric(`Sex`)) %>% 
  dplyr:: mutate(check = ifelse(`Sex`==`Training`, "NO ISSUE", "Sex <> Training")) %>%
    dplyr::filter(`check`== "Sex <> Training") %>%
    dplyr::mutate(`check_guidance`="Review the 'Sex'& 'Training' columns") %>% 
  dplyr::mutate(`check_num`="check11") %>% 
  dplyr::rename(uniqueid= id11_2)

#Dataset to check if Sex <> Work Cadre
check12<- df2 %>% 
  dplyr::filter(`Reporting Date` == "2022-07-31")%>%  #UPDATE DATE
  dplyr::filter(`Dataelement Code`==  "CV.2.4-18") %>% 
  tidyr::unnest(Value) %>% 
  dplyr::mutate(`Value`= as.numeric(`Value`)) %>%
  dplyr::mutate(id12=paste(Partner, OrgUnit, `Dataelement Code`, `Reporting Date`,`disagg category`, sep= "_")) %>% 
  dplyr::select(-c(Disaggregate,uniqueid)) %>%
  dplyr::mutate_at(vars("Value"), ~replace_na(.,0)) %>% 
  dplyr::mutate(`Value`= as.numeric(`Value`)) %>%
  dplyr::group_by(id12) %>% 
  dplyr::mutate(`Value`= sum(`Value`)) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(id12_2=paste(Partner, OrgUnit, `Dataelement Code`, `Reporting Date`, sep= "_")) %>%
  dplyr::select(-c(id12)) %>% 
  dplyr::distinct() %>% 
  tidyr::spread(key=`disagg category`, value=Value) %>% 
  dplyr::mutate_at(vars("Work cadre"), ~replace_na(.,0)) %>% 
  dplyr::mutate_at(vars("Sex"), ~replace_na(.,0)) %>% 
  dplyr::mutate(`Work cadre`= as.numeric(`Work cadre`)) %>% 
  dplyr::mutate(`Sex`= as.numeric(`Sex`)) %>% 
  dplyr:: mutate(check = ifelse(`Sex`==`Work cadre`, "NO ISSUE", "Sex <> Work cadre")) %>%
  dplyr::filter(`check`== "Sex <> Work cadre") %>%
  dplyr::mutate(`check_guidance`="Review the 'Sex'& 'Work cadre' columns") %>% 
  dplyr::mutate(`check_num`="check12") %>% 
  dplyr::rename(uniqueid= id12_2)

#Dataset to check if Severity of event <>Type of USG support
check13<- df2 %>% 
  dplyr::filter(`Reporting Date` == "2022-07-31")%>%  #UPDATE DATE
  dplyr::filter(`Dataelement Code`==  "CV.1.5-9") %>% 
  tidyr::unnest(Value) %>% 
  dplyr::mutate(`Value`= as.numeric(`Value`)) %>%
  dplyr::mutate(id13=paste(Partner, OrgUnit, `Dataelement Code`, `Reporting Date`,`disagg category`, sep= "_")) %>% 
  dplyr::select(-c(Disaggregate,uniqueid)) %>%
  dplyr::mutate_at(vars("Value"), ~replace_na(.,0)) %>% 
  dplyr::mutate(`Value`= as.numeric(`Value`)) %>%
  dplyr::group_by(id13) %>% 
  dplyr::mutate(`Value`= sum(`Value`)) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(id13_2=paste(Partner, OrgUnit, `Dataelement Code`, `Reporting Date`, sep= "_")) %>%
  dplyr::select(-c(id13)) %>% 
  dplyr::distinct() %>% 
  tidyr::spread(key=`disagg category`, value=Value) %>% 
  dplyr::mutate_at(vars("Severity of event"), ~replace_na(.,0)) %>% 
  dplyr::mutate_at(vars("Type of USG support"), ~replace_na(.,0)) %>% 
  dplyr::mutate(`Severity of event`= as.numeric(`Severity of event`)) %>% 
  dplyr::mutate(`Type of USG support`= as.numeric(`Type of USG support`)) %>% 
  dplyr:: mutate(check = ifelse(`Type of USG support`==`Severity of event`, "NO ISSUE", "Type of USG support <> Severity of event")) %>%
 dplyr::filter(`check`== "Type of USG support <> Severity of event") %>%
  dplyr::mutate(`check_guidance`="Review the 'Type of USG support'& 'Severity of event' columns") %>% 
  dplyr::mutate(`check_num`="check13") %>% 
  dplyr::rename(uniqueid= id13_2)

#Bind level 2 checks into 1 dataset
Level2 <- bind_rows(check3, check4, check5, check6, check8, check9a, check10, check11,check12, check13) %>% 
  dplyr::select('Partner',	'OrgUnit',	'Data element',	'Dataelement Code',	'Original Dataelement Code',	'NumDenom',	'disagg category',	'Disaggregate',	'Start Date',	'End date',	'Indicator Status',	'Reporting Date',		'uniqueid',	'check',	'check_guidance',	'check_num',	'Sex',	'Vaccine brand',	'Value',	'value_prev_month',	'Percent_change',	'Training',	'Work cadre',	'Severity of event',	'Type of USG support') #MUST be in this order so it appends to the googlesheet.

```

##Step 8: Create separate partner files by level

```{r}
  Level1_ANOVA<-Level1 %>% 
    dplyr::filter(`Partner`== "ANOVA")
  Level2_ANOVA<-Level2 %>% 
    dplyr::filter(`Partner`== "ANOVA")

  Level1_GUIDEHOUSE<-Level1 %>% 
    dplyr::filter(`Partner`== "Guidehouse")
  Level2_GUIDEHOUSE<-Level2 %>% 
    dplyr::filter(`Partner`== "Guidehouse")
  
  Level1_GETF_NEXTMILE<-Level1 %>% 
    dplyr::filter(`Partner`== "GETF")
  Level2_GETF_NEXTMILE<-Level2 %>% 
    dplyr::filter(`Partner`== "GETF")
  
   Level1_ADAPT<-Level1 %>% 
    dplyr::filter(`Partner`== "Right to Care" |`Partner`== "THE AURUM INSTITUTE"|`Partner`== "Foundation for Professional Development (FPD)"|`Partner`== "Genesis Analytics"|`Partner`== "Institute of Health Programs and Systems (IHPS)"|`Partner`== "IntraHealth"|`Partner`== "Qode Health Solutions"|`Partner`== "Re-Action!"|`Partner`== "UNIVERSITEIT STELLENBOSCH")
   Level2_ADAPT<-Level2 %>% 
        dplyr::filter(`Partner`== "Right to Care" |`Partner`== "THE AURUM INSTITUTE"|`Partner`== "Foundation for Professional Development (FPD)"|`Partner`== "Genesis Analytics"|`Partner`== "Institute of Health Programs and Systems (IHPS)"|`Partner`== "IntraHealth"|`Partner`== "Qode Health Solutions"|`Partner`== "Re-Action!"|`Partner`== "UNIVERSITEIT STELLENBOSCH")
```

##Step 9a: PRECLEAN PERIOD ONLY: Automatically append PRECLEAN results to existing partner DQRT google spreadsheets. Note that if you rerun it mutliple tabs, it will update in the DQRT spreadsheets multiple times.
```{r}

ANOVA_url_level1 <- "https://docs.google.com/spreadsheets/d/15fB-CoN_etdy9YOZ_IhGL8mePexudBj1eRXg1VrDho0/edit#gid=224915941"
  ss_ANOVA_level1<-drive_get(ANOVA_url_level1)
  ss_ANOVA_level1 %>% sheet_append(Level1_ANOVA, sheet = "Level1")

ANOVA_url_level2 <- "https://docs.google.com/spreadsheets/d/15fB-CoN_etdy9YOZ_IhGL8mePexudBj1eRXg1VrDho0/edit#gid=210686588"
  ss_ANOVA_level2<-drive_get(ANOVA_url_level2)
  ss_ANOVA_level2 %>% sheet_append(Level2_ANOVA, sheet = "Level2")

GUIDEHOUSE_url_level1 <- "https://docs.google.com/spreadsheets/d/16IXhexNBBboJNygo26-ImtkFvsLxzw6_loUQvzy4Q5A/edit#gid=1902024796"
  ss_GUIDEHOUSE_level1<-drive_get(GUIDEHOUSE_url_level1)
  ss_GUIDEHOUSE_level1 %>% sheet_append(Level1_GUIDEHOUSE, sheet = "Level1")

GUIDEHOUSE_url_level2 <- "https://docs.google.com/spreadsheets/d/16IXhexNBBboJNygo26-ImtkFvsLxzw6_loUQvzy4Q5A/edit#gid=1079651008"
  ss_GUIDEHOUSE_level2<-drive_get(GUIDEHOUSE_url_level2)
  ss_GUIDEHOUSE_level2 %>% sheet_append(Level2_GUIDEHOUSE, sheet = "Level2")

  GETF_NEXTMILE_url_level1 <- "https://docs.google.com/spreadsheets/d/1n_8PWl1cUKGi-8aACrsJW4GbcNAan7zvqEqdIyxmaG4/edit#gid=737504841"
  ss_GETF_NEXTMILE_level1<-drive_get(GETF_NEXTMILE_url_level1)
  ss_GETF_NEXTMILE_level1 %>% sheet_append(Level1_GETF_NEXTMILE, sheet = "Level1")

GETF_NEXTMILE_url_level2 <- "https://docs.google.com/spreadsheets/d/1n_8PWl1cUKGi-8aACrsJW4GbcNAan7zvqEqdIyxmaG4/edit#gid=1986328981"
  ss_GETF_NEXTMILE_level2<-drive_get(GETF_NEXTMILE_url_level2)
  ss_GETF_NEXTMILE_level2 %>% sheet_append(Level2_GETF_NEXTMILE, sheet = "Level2")

    ADAPT_url_level1 <- "https://docs.google.com/spreadsheets/d/1kodgE7uAQs91CDEgCPzKaV4KAkMZIDRtpaj4m2A3-Og/edit#gid=1471368329"
  ss_ADAPT_level1<-drive_get(ADAPT_url_level1)
  ss_ADAPT_level1 %>% sheet_append(Level1_ADAPT, sheet = "Level1")

ADAPT_url_level2 <- "https://docs.google.com/spreadsheets/d/1kodgE7uAQs91CDEgCPzKaV4KAkMZIDRtpaj4m2A3-Og/edit#gid=2060148287"
  ss_ADAPT_level2<-drive_get(ADAPT_url_level2)
  ss_ADAPT_level2 %>% sheet_append(Level2_ADAPT, sheet = "Level2")

  
```
##Step 9b: CLEAN PERIOD ONLY: Automatically append POSTCLEAN results to master  DQRT google spreadsheet; ONLY RUN THIS AFTER PARTNER DATA HAS BEEN REVIEWED AND UPDATES HAVE BEEN MADE TO TOOLS BY PARTNERS.Note that if you rerun it mutliple tabs, it will update in the DQRT spreadsheets multiple times.

```{r}
master_url_level1 <- "https://docs.google.com/spreadsheets/d/1YkvbqRCqGA4LAE3YjbSC8jzaN64rZ3RhbzfpA1yrdLU/edit#gid=1681946868"
ss_master_level1<-drive_get(master_url_level1)
ss_master_level1 %>% sheet_append(Level1, sheet = "Level1")

master_url_level2 <-"https://docs.google.com/spreadsheets/d/1YkvbqRCqGA4LAE3YjbSC8jzaN64rZ3RhbzfpA1yrdLU/edit#gid=1492698209"
ss_master_level2<-drive_get(master_url_level2)
ss_master_level2 %>% sheet_append(Level2, sheet = "Level2")

#After running this, check master file to make sure the new data appended in the tabs.
#If you run it mutliple times, the flags will append to the google sheet multiple times.

```
##Step 9c OPTIONAL/IF ISSUE WITH A and B: Save results in existing Excel Workbook on desktop

```{r}
# https://ycphs.github.io/openxlsx/articles/Introduction.html
  
ANOVA_wb<- openxlsx::loadWorkbook("ANOVA_COVID-19 DQRT Tracker.xlsx") 

  openxlsx::writeData(ANOVA_wb, sheet = "Level1", x = Level1_ANOVA, 
          colNames=T, withFilter=T)
  openxlsx::writeData(ANOVA_wb, sheet = "Level2", x = Level2_ANOVA, 
          colNames=T, withFilter=T)
  openxlsx::saveWorkbook(ANOVA_wb, "ANOVA_COVID-19 DQRT Tracker.xlsx", overwrite = TRUE)
  
GUIDEHOUSE_wb<- openxlsx::loadWorkbook("GUIDEHOUSE_COVID-19 DQRT Tracker.xlsx") 
  openxlsx::writeData(GUIDEHOUSE_wb, sheet = "Level1", 
  x = Level1_GUIDEHOUSE, 
  colNames=T, withFilter=T)
  openxlsx::writeData(GUIDEHOUSE_wb, sheet = "Level2", 
  x = Level2_GUIDEHOUSE, 
  colNames=T, withFilter=T)
  openxlsx::saveWorkbook(GUIDEHOUSE_wb, "GUIDEHOUSE_COVID-19 DQRT Tracker.xlsx", overwrite = TRUE)

GETF_NETXTMILE_wb<- openxlsx::loadWorkbook("GETF_COVID-19 DQRT Tracker.xlsx")
  openxlsx::writeData(GETF_NETXTMILE_wb, sheet = "Level1", 
  x =   Level1_GETF_NEXTMILE, 
  colNames=T, withFilter=T)
  openxlsx::writeData(GETF_NETXTMILE_wb, sheet = "Level2", 
  x =   Level2_GETF_NEXTMILE, 
  colNames=T, withFilter=T)
  openxlsx::saveWorkbook(GETF_NETXTMILE_wb, "GETF_COVID-19 DQRT Tracker.xlsx", overwrite = TRUE)
  
ADAPT_wb<- openxlsx::loadWorkbook("ADAPT_COVID-19 DQRT Tracker.xlsx") 
  openxlsx::writeData(ADAPT_wb, sheet = "Level1", 
  x =  Level1_ADAPT, 
  colNames=T, withFilter=T)
  openxlsx::writeData(ADAPT_wb, sheet = "Level2", 
  x =  Level2_ADAPT, 
  colNames=T, withFilter=T)
  openxlsx::saveWorkbook( ADAPT_wb, "ADAPT_COVID-19 DQRT Tracker.xlsx", overwrite = TRUE)

##If need a master file-optional to run
  wb<- openxlsx::loadWorkbook("MASTER_COVID-19 DQRT Tracker.xlsx") 
  openxlsx::writeData(wb, sheet = "Level1", x = Level1, 
          colNames=T, withFilter=T)
  openxlsx::writeData(wb, sheet = "Level2", x = Level2, 
          colNames=T, withFilter=T) 
  openxlsx::saveWorkbook(wb, "MASTER_COVID-19 DQRT Tracker.xlsx", overwrite = TRUE)
  

##If needed To output to master CSV file directly
#write_excel_csv(Level1, paste0("SA_COVID19_LEVEL1_DataVal_DataQual_", format(Sys.time(), "%d-%b-%Y"), ".csv"))
#write_excel_csv(Level2, paste0("SA_COVID19_LEVEL2_DataVal_DataQual_", format(Sys.time(), "%d-%b-%Y"), ".csv"))

```
